---
title: "data624_project_1"
author: "Richard"
date: "4/2/2022"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
library(tidyverse)
library(fpp3)
library(readxl)
```

# ATM Forecast

```{r}
# function selects chosen atm
select_atm = function(n){
  atm = paste0('ATM',n)
  data = atm_series%>%
    filter(ATM == atm)
  data = data%>%
    replace_na(list(Cash = mean(data$Cash,na.rm=TRUE)))
  return(data)
}

# plots time series for chosen ATM
plot_atm = function(n){
  data = select_atm(n)
  data%>%
    autoplot(Cash)+labs(title = paste('Withdrawals from',paste0('ATM',n)))
}



# plots forecasts based on model
forecast_model = function(data, model,type){
  model%>%
  select('ATM',type)%>%
  forecast(h=30)%>%
  autoplot(data)+labs(title = paste(data$ATM[1],'Withdrawal Forecast'))
}
```


```{r}
atm_data = read_excel('ATM624Data.xlsx')
atm_data%>%head()
```

#### cleaning data

```{r}
# transform excel date values to dates
atm_data['DATE'] = as.Date(atm_data$DATE, origin = "1899-12-30")

# turn dataset into time series
atm_series = atm_data%>%
  as_tsibble(index = DATE, key = ATM)

atm_series%>%head()
```

#### ATM1

ATM withdrawals seem to follow a flat(not increasing nor decreasing) trend with seasonal variation

```{r}
atm1 = select_atm(1)
plot_atm(1)
```

Using classical decomposition, we can see that there is no clear trend. However, there is a seasonal component

```{r}
atm1%>%
  model(
    classical_decomposition(Cash)
    )%>%
  components()%>%
  autoplot()
```

Chose various models to fit the data. Based on RMSE, the arima model performed the best for ATM1

```{r}
atm1.models = atm1%>%
  model(
    'mean' = MEAN(Cash),
    'snaive' = SNAIVE(Cash),
    'ets' = ETS(Cash~error("A") + trend("N") + season("A")),
    'arima' = ARIMA(Cash~pdq(1,0,1))
  )

atm1.models%>%
  accuracy()
```

Residuals of the arima model are normally distributed and uncorrelated

```{r}
atm1.models%>%
  select('arima')%>%
  gg_tsresiduals()
```

Using arima model to forecast may 2010

```{r}
forecast_model(atm1,atm1.models,'arima')

```


#### ATM2

Withdrawals from ATM2 are similar to those seen in ATM1. There seems to be a seasonal component and no clear trend

```{r}
atm2 = select_atm(2)
plot_atm(2)
```

Based on the decomposition, we see that there is a slight downtrend. 

```{r}
atm2%>%
  model(
    classical_decomposition(Cash)
    )%>%
  components()%>%
  autoplot()
```

Fitting various models and calculating the RMSE, we can see that arima model slightly out performs the ETS model

```{r}
atm2.models = atm2%>%
  model(
    'mean' = MEAN(Cash),
    'snaive' = SNAIVE(Cash),
    'ets' = ETS(Cash~error("A") + trend("A") + season("A")),
    'arima' = ARIMA(Cash~pdq(1,0,1))
  )

atm2.models%>%
  accuracy()
```

Residuals for the arima model are normal and uncorrelated

```{r}
atm2.models%>%
  select('arima')%>%
  gg_tsresiduals()
```

Using Arima model, forecast ATM2 withdrawals

```{r}
forecast_model(atm2,atm2.models,'arima')
```



#### ATM3

ATM3 only has withdrawals data for a couple of days in the dataset.There is no seasonal component

```{r}
atm3 = select_atm(3)
plot_atm(3)
```

comparing various models, we can see that the naive method is nearly as good as ETS and ARIMA while being far less complex

```{r}
atm3.models = atm3%>%
  model(
    'mean' = MEAN(Cash),
    'naive' = NAIVE(Cash),
    'ets' = ETS(Cash~error("A") + trend("A") + season("N")),
    'arima' = ARIMA(Cash~pdq(1,1,0))
  )

atm3.models%>%
  accuracy()
```

Since the dataset has so few entries, I believe using the naive method is superior as there is little historical data to go off of. 

```{r}
forecast_model(atm3,atm3.models,'naive')
```



#### ATM4

ATM4 follows a similar pattern to ATM 1 and ATM2. There seems to be a seasonal component with no general trend. There also seems to be a couple of days in particular with abnormally high amounts of cash withdrawal

```{r}
atm4 = select_atm(4)
plot_atm(4)
```

After replacing outliers with the mean, we are able to see the fluctuations more clearly.

```{r}

m = mean(atm4$Cash)
s = sd(atm4$Cash)
low = m-3*s
high = m+3*s

atm4 = atm4%>%
  filter(Cash>=low & Cash<= high)%>%
  fill_gaps()%>%
  replace_na(list(Cash = m))

atm4%>%
  autoplot(Cash)+labs(title = 'Withdrawals from ATM4')

```

There is a seasonal component and a flat trend

```{r}
atm4%>%
  model(
    classical_decomposition(Cash)
    )%>%
  components()%>%
  autoplot()
```

All of the models have high RMSE values. Our ets model performs the best with lowest RMSE

```{r}
atm4.models = atm4%>%
  model(
    'mean' = MEAN(Cash),
    'snaive' = SNAIVE(Cash),
    'ets' = ETS(Cash~error("A") + trend("N") + season("A")),
    'arima' = ARIMA(Cash~pdq(1,0,1))
  )

atm4.models%>%
  accuracy()
```
The residuals are near normal and uncorrelated

```{r}
atm4.models%>%
  select('ets')%>%
  gg_tsresiduals()
```

Forecasting ATM4 using ETS model

```{r}
forecast_model(atm4,atm4.models,'ets')
```


# Forecasting Power Usage

```{r}
power_usage = read_excel('ResidentialCustomerForecastLoad-624.xlsx')
power_usage%>%head()
```

#### Cleaning data

```{r}
power_usage['yearmonth'] = yearmonth(power_usage$`YYYY-MMM`)

# get mean, sd and create low/high based on 3 standard deviations
KWH_mean = mean(power_usage$KWH, na.rm=TRUE)
KWH_sd = sd(power_usage$KWH, na.rm=TRUE)
KWH_low = KWH_mean-KWH_sd*3
KWH_high = KWH_mean+KWH_sd*3

# replace NA with mean
power_series = power_usage%>%
  select('yearmonth','KWH')%>%
  as_tsibble(index = 'yearmonth')

power_series%>%head()

```

Data looks to be seasonal with an flat trend. There seems to be an outlier near 2010

```{r}
power_series%>%
  autoplot(KWH) + labs(title = 'KWH')
```

remove outliers from dataset and fill gaps with mean

```{r}
power_usage_ts = power_series%>%
  filter(KWH>= KWH_low & KWH<=KWH_high)%>%
  fill_gaps()%>%
  replace_na(list(KWH = KWH_mean))

power_usage_ts%>%
  autoplot(KWH) + labs(title = 'KWH')
  
```


#### Decomposition

There is a seasonal component along with trend. There is a slight uptrend

```{r}
power_usage_ts%>%
  model(
    classical_decomposition(KWH)
    )%>%
  components()%>%
  autoplot()
```

#### Modeling

Based on RMSE, the arima model fits this dataset the best

```{r}
power.models = power_usage_ts%>%
  model(
    'mean' = MEAN(KWH),
    'snaive' = SNAIVE(KWH),
    'ets' = ETS(KWH~error("A") + trend("A") + season("A")),
    'arima' = ARIMA(KWH~pdq(2,1,2) + PDQ(0,1,2))
  )

power.models%>%
  accuracy()
```

Model residuals are normal and uncorrelated

```{r}
power.models%>%
  select('arima')%>%
  gg_tsresiduals()
```


#### Forecast

Forecast KWH usage using arima

```{r}
power.models%>%
  select('arima')%>%
  forecast(h=30)%>%
  autoplot(power_usage_ts) + labs(title ='KWH Forecast')

```



# Waterflow Pipe (Bonus)

```{r}
get_lambda = function(data, feature){
  data%>%
    features(data[feature],features = guerrero)%>%
    pull(lambda_guerrero)%>%
    return()
}
```


```{r}
water_pipe1 = read_excel('Waterflow_Pipe1.xlsx')
water_pipe2 = read_excel('Waterflow_Pipe2.xlsx')
water_pipe1%>%head()
water_pipe2%>%head()
```

#### cleaning data

```{r}
water_pipe = rbind(water_pipe1,water_pipe2)

mean_flow = mean(water_pipe$WaterFlow,na.rm =TRUE)

water_series = water_pipe%>%
  mutate(days = as.Date(water_pipe$`Date Time`,origin = "1899-12-30"),
         hour = round((water_pipe$`Date Time`- round(water_pipe$`Date Time`))*24))%>%
  mutate(day_hour = days+hours(hour))%>%
  select('day_hour','WaterFlow')%>%
  group_by(day_hour)%>%
  summarize(avg_flow = mean(WaterFlow))%>%
  as_tsibble(index = day_hour)%>%
  fill_gaps()%>%
  replace_na(list(avg_flow = mean_flow))


# Average flow increased substantially in November
water_series%>%
  autoplot(avg_flow) +labs(title = 'Average Water Flow')
```

#### Exploratory Analysis

```{r}
water_series%>%
  model(
    classical_decomposition(avg_flow)
    )%>%
  components()%>%
  autoplot()
```

After differencing the data fits better between the confidence levels

```{r}

water_series%>%
  ACF(avg_flow)%>%
  autoplot() + labs(title = 'ACF Flow')

water_series%>%
  ACF(difference(avg_flow))%>%
  autoplot() + labs(title = 'ACF Difference in Flow')
```

#### Modeling

Based on the RMSE, The ARIMA model hasthe least amount of error.

```{r}
flow.models = water_series%>%
  model(
    'mean' = MEAN(avg_flow),
    'naive' = NAIVE(avg_flow),
    'snaive' = SNAIVE(avg_flow),
    'ets' = ETS(avg_flow~error("A") + trend("N") + season("A")),
    'arima' = ARIMA(avg_flow~pdq(2,1,2) + PDQ(0,1,1))
  )

flow.models%>%
  accuracy()
```

```{r}
flow.models%>%
  select('arima')%>%
  gg_tsresiduals()
```

#### Forecast

```{r}
flow.models%>%
  select('arima')%>%
  forecast(h=24*7)%>%
  autoplot(water_series)+labs(title = 'Average Water Flow Forecast')

```



