---
title: "DATA607 HW3"
author: "Richard"
date: "2/16/2021"
output: html_document
---
packages used
```{r}
library(tidyverse)
```

# 1)

```{r}
majors = read.csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/college-majors/majors-list.csv", header= TRUE, sep=",")

data_or_stat = function(arr){
  list = c()
  for (major in arr){
    if(grepl('data',tolower(major)) | grepl('statistics',tolower(major))){
      list = append(list,major)
    }
  }
  return(list)
}


major_list = majors$Major
data_or_stat(major_list)


#found an easier way to do it in R after writing out my solution XD
grep('data|statistics', major_list,value=TRUE,ignore.case=TRUE)


```
# 2)

```{r}
fruits= c("bell pepper", "bilberry", "blackberry", "blood orange", "blueberry", "cantaloupe", "chili pepper", "cloudberry", "elderberry", "lime", "lychee", "mulberry", "olive", "salal berry")
string = "c('"
for (fruit in fruits){
  if(fruit == fruits[1]){
    string = paste(string,fruit,"'",sep='')
  }
  else{
    fruit = paste("'", fruit,"'",sep = "")
    string = paste(string,fruit, sep=", ")
    }
}
string = paste(string,')')
string

```
# 3) Define Regex Expressions:
<pre>Note: R requires \\1 to reference first call group while regex syntax is \1 </pre>

### 1: (.)\1\1   

matches a string that contains a character repeated 3 times in a row (except for line breaks) (ex:aaa)

```{r}
strings = c('abc','aaa','bbb','xyz')
pattern = '(.)\\1\\1'
str_detect(strings,pattern)
```


### 2: "(.)(.)\\2\\1"  
matches a string that contains any two characters followed those two characters reversed (ex: ahha)

```{r}
strings = c('abcd','aaaa','ahha','haha')
pattern = "(.)(.)\\2\\1"
str_detect(strings,pattern)
```


### 3: (..)\1          
matches a string that contains any two characters repeated twice (ex:ezez)

```{r}
strings = c('ezez','aaaa','haha','aboab')
pattern = "(..)\\1"
str_detect(strings, pattern)
```


### 4: "(.).\\1.\\1" 
matches a string that contains three characters repeated (ex: abacad)

```{r}
strings = c('abacad','aabbcc','three','aaa')
pattern = "(.).\\1.\\1"
str_detect(strings,pattern)
```




### 5: "(.)(.)(.).*\\3\\2\\1"

matches a string that contains any three characters and its reverse (ex: abcdcba)

```{r}
strings = c('cccccc','abcdcba','abcdabc','ahaha')
pattern = "(.)(.)(.).*\\3\\2\\1"
str_detect(strings,pattern)
```


# 4) Create Regex Expressions:

### 1) Start and end with the same character.
 <pre>^([a-z])[a-z]*\1$</pre>
```{r}

names = c('richard','hannah','bob','annie')
pattern = '^([a-z])[a-z]*\\1$'
str_detect(names, pattern)


```


### 2) Contain a repeated pair of letters (e.g. "church" contains "ch" repeated twice.)
 <pre>[a-z]*([a-z]{2})[a-z]*\1[a-z]*</pre>
 
```{r}
words = c('church','dog','halal','cat')
pattern = "[a-z]*([a-z]{2})[a-z]*\\1[a-z]*"
str_detect(words,pattern)
```
 

### 3) Contain one letter repeated in at least three places (e.g. "eleven" contains three "e"s.)

<pre>[a-z]*([a-z])[a-z]*\1[a-z]*\1[a-z]*</pre>

```{r}
numbers = c('eleven','seven','three','seventeen')
pattern = '[a-z]*([a-z])[a-z]*\\1[a-z]*\\1[a-z]*'
str_detect(numbers, pattern)
```

  