---
title: "DATA607 hw1"
author: "Richard"
date: "2/7/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Exploratory data analysis on variables that may affect why people don't vote

Loading data into dataframe 
```{r}
non_voter = read.csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/non-voters/nonvoters_data.csv", header= TRUE, sep=",")
```
Creating a subset,renaming columns and droping na values
```{r}
subset_voter = subset(non_voter,select = c(ppage,educ,race,gender,income_cat))

colnames(subset_voter)=c("age","education","race","gender","income")
subset_voter = na.omit(subset_voter)

head(subset_voter)
```

```{r}
hist(subset_voter$age)
```
Distribution for age of non-voters
```{r}
income_count = table(subset_voter$income)
barplot(income_count)
```
Bar graph for the incomes of non-voters

```{r}
education_count = table(subset_voter$education)
barplot(education_count)
```
Bar graph for education of non-voters
```{r}
gender_count = table(subset_voter$gender)
barplot(gender_count)
```
Gender count of non-voters


Conclusion:
1:The distribution for age of non voters is not normally distributed and there seem to be two peaks: 1) Late 20;s and 2) In the 60's
2: People in any income group are equally likely to not vote except for the $75-125k group. They are slightly more likely to not vote.
3: People who are more educated seem more likely to not vote as there are significantly more college educated people in this sample
4: Men and women are both equally likely to not vote.
