---
title: "data607 hw5"
author: "Richard"
date: "3/7/2021"
output: html_document
---
# Packages 

```{r}
library(tidyverse)
```

# Loading Dataset

```{r}
url = "https://raw.githubusercontent.com/schoolkidrich/R/main/DATA%20607/week5/delay.csv"

delays = read.csv(url)
head(delays)
```

# pivoting dataset on cities

```{r}
len = dim(delays)[2]
col = names(delays)
cities = pivot_longer(delays,col[3:len],names_to = 'city')
head(cities)
```

# pivoting dataset on tardiness

```{r}
clean_data = pivot_wider(cities,names_from = "Tardiness")
names(clean_data)[3] = 'on_time'
head(clean_data)


```

# on time by cities (percent)

#### Alaksa airlines has a higher rate of arriving on time for every city

```{r}
clean_data$percent_on_time = clean_data$on_time/(clean_data$on_time+clean_data$delayed)

ggplot(clean_data,aes(x = reorder(city,-percent_on_time), y = percent_on_time, fill = Airline)) + geom_bar(stat='identity', position = position_dodge()) + labs(x = "city",title = "Percentage of Timely Arrivals by City")
```

# by count

#### There are more on time fligts for Alaska in some cities and more for Am west in others
```{r}
ggplot(clean_data,aes(x = reorder(city,-on_time), y = on_time, fill = Airline)) + geom_bar(stat='identity', position = position_dodge()) + labs(x = "city", title = "Count of Timely Arrivals by City")
```

# on time by airlines (percent)

#### unsurprisingly, Alaska airline also has a higher rate of being on time in total

```{r}
airline_on_time = group_by(clean_data,Airline) %>%
  summarize(percent = sum(on_time)/(sum(on_time)+sum(delayed)))

ggplot(airline_on_time, aes(x = Airline, y = percent))+geom_bar(stat='identity') + labs(title = "Percentage of Timely Arrivals")

```

# by count

#### There are more on time flights in total from AM WEST. This probably due to the abundance of flights in the city of pheonix where flights are primarily AM WEST. 

```{r}
ggplot(clean_data, aes(x = Airline,y=on_time))+geom_bar(stat="identity")+
  labs(title = "Count of Timely Arrivals")
```

# Conclusion

#### Although Alaksa airlines out performs AM WEST, in terms of tardiness rate, in every city the overall performance is worse. This is because most of AM WESTS flights are from pheonix while most of Alaskas flights are in Seattle. AM WEST in pheonix outperforms Alaska in Seattle which and those are the cities that are weighted the most (simpsons paradox)  
