---
title: "data607 project 2"
author: "Richard"
date: "3/14/2021"
output: html_document
---

# Loading Packages

```{r setup, include=TRUE}
library(tidyverse)
```


# Untidy datasets

```{r}

squirrel_url = "https://raw.githubusercontent.com/schoolkidrich/R/main/DATA%20607/project2/2018_Central_Park_Squirrel_Census_-_Squirrel_Data.csv"

gdp_url = "https://raw.githubusercontent.com/schoolkidrich/R/main/DATA%20607/project2/gdp.csv"

candy_url = "https://raw.githubusercontent.com/schoolkidrich/R/main/DATA%20607/project2/candyhierarchy2017.csv"

```

# 1) Squirrel dataset

```{r}

squirrel_df = read.csv(squirrel_url)
head(squirrel_df)

```

#### There are a lot of unesseccary data and seperate rows for various actions. I will drop columns I think are unessecary and pivot the action colu1ms into one "Action" column

```{r}
# dropping columns
squirrel_sightings = squirrel_df[c('Unique.Squirrel.ID','Hectare','Shift','Age','Primary.Fur.Color','Running','Chasing','Climbing','Eating','Foraging')]

# pivoting dataframe
squirrel_sightings = squirrel_sightings %>%
pivot_longer(names(squirrel_sightings)[6:dim(squirrel_sightings)[2]], names_to = "Action")

# dropping FALSE (N/A) values
squirrel_clean = squirrel_sightings[squirrel_sightings['value'] == TRUE,]
squirrel_clean = squirrel_clean[names(squirrel_clean)[1:dim(squirrel_clean)[2]-1]]

head(squirrel_clean)

```

#### What are squirrels at Central Park up to?

```{r}
squirrel_clean %>%
  group_by(Action) %>%
  summarise(Percentage = n()/dim(squirrel_clean)[1]) %>%
  ggplot(mapping = aes(x = reorder(Action,-Percentage),y = Percentage)) +geom_bar(stat = 'identity') + labs(x = 'Actions', title = "What are Squirrels at Central Park up to?")
```

#### Which location has the most squirrels?
```{r}
squirrel_clean %>%
  group_by(Hectare) %>%
  summarise(Count = n()) %>%
  arrange(desc(Count))%>%
  head(10) %>%
  ggplot(mapping = aes(x = reorder(Hectare,-Count),y = Count))+geom_bar(stat = 'identity') + labs(x = 'Location', title = 'Locations with the Most Squirrel Sightings')
```

# Conclusion:

#### Squirrels at central park are usually either looking for or eating food. If you want to see squirrels I suggest you go to location 14E as most of the sightings are from there 


# 2) GDP dataset

```{r}
gdp_df = read.csv(gdp_url)
head(gdp_df)
```

#### I want to remove rename columns to remove the "X" and then pivot all the year columns into one 'Year' column. Afterwards I will remove NA values

```{r}
# removing "x" from year strings
cols = names(gdp_df)
len = dim(gdp_df)[2]
for (col in cols[5:len]){
  cols[cols == col] = substr(col,2,5) 
}
names(gdp_df) = cols

# pivoting year columns to a single 'Year' column
gdp_clean = gdp_df %>%
  pivot_longer(cols = cols[5:len],names_to = "Year", values_to = 'percent_growth') %>%
  drop_na()

# separate df for country codes
country_code = unique(gdp_clean[c('Country.Name','Country.Code')])

# drop unwanted columns from gdp df

gdp_clean = gdp_clean[c('Country.Name','Year','percent_growth')]

head(gdp_clean)

```

#### Country with the highest change in GDP

```{r}
gdp_clean %>%
  group_by(Country.Name) %>%
  summarize(Avg_Growth = sum(percent_growth)/n())%>%
  arrange(desc(Avg_Growth)) %>%
  head(10) %>%
  ggplot(mapping = aes(x = reorder(Country.Name,Avg_Growth),y = Avg_Growth))+geom_bar(stat = 'identity') + coord_flip() +labs(x = 'Countries', title = 'Largest Average Change in GDP (1960-2020)')
                    
```

#### Change in GDP (United States) since I was born

```{r}
# subset df into united states only
us_gdp = gdp_clean[gdp_clean['Country.Name'] == 'United States' & gdp_clean['Year']>= 1997,]

us_gdp  %>%
  ggplot(mapping = aes(x = Year, y = percent_growth, group = Country.Name))+geom_line() + theme(axis.text.x = element_text(angle = 90)) + labs(title = 'Change in GDP since 1997')

```

# Colclusion
#### The country with the greatest average growth in GDP is Guinea. This was surprising to me especially since other countries such as China and Japan have grown so much but I believe its because they are a relatively poorer country which means higher potential growth (%) wise. 

#### For the United States, I noticed that growth was more volatile prior to 2011. I also notice a sharp decline in growth leading into 2009 (Great Recession)

# 3) Candy Dataset

```{r}
candy_df = read.csv(candy_url)
head(candy_df)

```

#### Noticed that there are a lot of characters such as "Q#" and "." in column names and I want to remove those

```{r}
length = dim(candy_df)[2]
columns = names(candy_df)
pattern = '^[a-z][0-9]$'

# removing "Q#" from strings
for (column in columns){
  new_str = ""
  for (item in strsplit(column,"\\.")[[1]]){
    if (grepl(pattern, item, ignore.case = TRUE) == FALSE){
      new_str = paste(new_str,item)
    }
  }
  columns[columns == column] = trimws(new_str)
}
names(candy_df) = columns
head(candy_df)
```

#### Similar to the last dataset this one is very wide and I want to pivot all the candy columns into one with a rating column that has the values: "MEH", "JOY", "DESPAIR"

```{r}
# pivoting candy data
candy_clean = candy_df %>%
  pivot_longer(columns[7:(length-11)],names_to = "Candy", values_to = "Mood")

# Separate out repeated information
candy_id = unique(candy_clean[c('Internal ID','GOING OUT','GENDER','AGE','COUNTRY')])

# Drop columns not needed for analysis
candy_clean = candy_clean[c('Internal ID','Candy','Mood')]
candy_clean = candy_clean[candy_clean["Mood"] != "",]
head(candy_clean)
```

#### Joy vs Despair

```{r}

candy_clean%>% 
  group_by(Mood)%>%
  summarise(Percent = n()/dim(candy_clean)[1]) %>%
  ggplot(mapping = aes(x = Mood, y = Percent)) + geom_bar(stat = 'identity') + labs(title = "")

```

#### Which candy do people like the most?

```{r}
joy = candy_clean[candy_clean['Mood'] == 'JOY',]
joy %>% 
  group_by(Candy)%>%
  summarize(Count = n())%>%
  arrange(desc(Count))%>%
  head(10) %>%
  ggplot(mapping = aes(x = reorder(Candy,Count),y = Count))+geom_bar(stat = 'identity') +coord_flip() + labs(x = 'Candy', title = "Which Candy Brings People Joy?")

  
```

# Conclusion

#### More candies bring people Joy than Despair... people just want a full sized candy bar. 
