---
title: "data 624 hw1"
author: "Richard"
date: "2/13/2022"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
library(tidyverse)
library(fpp3)
```

# 2.1

## a) autoplots
```{r}
autoplot(gafa_stock, Open)

PBS %>%
  filter(ATC2 == "A10")%>%
  autoplot(Cost)

autoplot(vic_elec, Temperature)

pelt%>%
  pivot_longer(c(Hare, Lynx))%>%
  autoplot(value)
  
```


## b)
```{r}
# historical stock prices (USD) and volume between 2014-2018 for goog, amzn, fb and aapl
# time interval is (trading) days
help(gafa_stock)


# monthly Australian medicare prescription data (number of prescriptions and costs(AUD) )
# time interval is monthly
help(PBS)


# Electricity demand for Victoria, Australia (Demand, temp, holiday indicator)
# time interval of 30 minutes
help(vic_elec)


# Annual fur pelt trading data from 1845 to 1935 (number of hare and lynx pelts traded)
# time interval is one year
help(pelt)
```


# 2.2

```{r}

# finds row where closing price is highest for a given ticker in gafa_stock dataset
max_close_price = function(ticker){
  stock = gafa_stock%>%
    filter(Symbol == ticker)%>%
    filter(Close == max(Close))%>%
    return()
}

# prints row where closing price is highest for each unique ticker in gafa_stock
tickers = gafa_stock$Symbol%>%
  unique()

for (ticker in tickers){
  print(max_close_price(ticker))
}

```

# 2.3


## a) load tute1

```{r}
tute1 = read.csv('https://raw.githubusercontent.com/schoolkidrich/CUNY_MSDS/main/DATA_624/hw1/tute1.csv')

head(tute1)
```

## b) convert data to time series

```{r}
tute1.series = tute1%>%
  mutate(Quarter = yearmonth(Quarter))%>%
  as_tsibble(index = Quarter)

head(tute1.series)
```

## c) construct time series plot


when you don't include facet_grid() all the lines are plotted onto a single graph
```{r}
tute1.series %>%
  pivot_longer(c(Sales,AdBudget,GDP))%>%
  ggplot(aes(x = Quarter, y = value, color = name))+
  geom_line()+ facet_grid(name ~., scales = "free_y")
```

# 2.4

## a) load USgas package

```{r}
library(USgas)
```

## b) create time series object from us_total with year and index and state as key

```{r}
us_total.series = us_total%>%
  as_tsibble(index = year, key= state)
  
head(us_total.series)
```

## c) plot plot annual gas consumption for new england area

```{r}
us_total.series%>%
  filter(state == c('Maine', 'Vermont', 'New Hampshire', 'Massachusetts', 'Connecticut', 'Rhode Island'))%>%
  ggplot(aes(x = year, y = y, color = state))+
  geom_line()+
  facet_grid(state ~., scale = 'free')+
  labs(title = 'Annual Gas Consumption for New England Area')

```

# 2.5

## a) read tourism.xlsx into R

```{r}
tour = readxl::read_excel('tourism.xlsx')
head(tour)
```

## b) create time series object using tour data

```{r}
tour.series = tour%>%
  mutate(Quarter = yearquarter(Quarter))%>%
  as_tsibble(index = Quarter, key = c(Region, Purpose))
head(tour.series)
```

## c) find combination of Region and Purpose that has maximum number of trips on average

```{r}
# summarize average trip length by region and purpose then finds combination with highest
tour%>%
  group_by(Region, Purpose)%>%
  summarize(avg_trip_length = mean(Trips), .groups = "keep")%>%
  ungroup()%>%
  filter(avg_trip_length == max(avg_trip_length))
```

## d) create new time series object which combines purposes and regions

```{r}
trips.state = tour.series%>%
  group_by(State)%>%
  summarize(total_trips = sum(Trips))

head(trips.state)
```

# 2.8

```{r}
set.seed(8009)
myseries <- aus_retail %>%
  filter(`Series ID` == sample(aus_retail$`Series ID`,1))

head(myseries)
```

overall, turnover has steadily increased over time but has plateaued since ~2010

```{r}
myseries%>%
  autoplot(Turnover)
```

Turnover seems to always be highest during November - December. This trend has always been the case, but in recent years the spike in turnover has increased. In recent years, there is also a dip in turnover around February and a smaller spike around may

```{r}
myseries%>%
  gg_season(Turnover)
```

