---
title: "data624_hw3"
author: "Richard"
date: "2/25/2022"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
library(tidyverse)
library(fpp3)
```

# 5.1

### Australian Population

```{r}
aus_pop = global_economy%>%
  filter(Country == 'Australia' & Year >= 2000)%>%
  select(Population)

aus_pop%>%
  model(RW(Population ~ drift()))%>%
  forecast()%>%
  autoplot(aus_pop)+
  labs(title = 'Australian Population Forecast')

```

### Bricks

```{r}
aus_brick = aus_production%>%
  na.omit()%>%
  select(Bricks)

aus_brick%>%
  model(NAIVE(Bricks))%>%
  forecast()%>%
  autoplot(aus_brick)+
  labs(title = 'Australian Brick Production Forecast')
```

### NSW Lambs

```{r}
nsw_lambs = aus_livestock%>%
  filter(Animal == 'Lambs' & State == 'New South Wales')%>%
  select(Count)

nsw_lambs%>%
  model(SNAIVE(Count))%>%
  forecast()%>%
  autoplot(nsw_lambs)+
  labs(title = 'Lamb Slaughter Count in New South Wales')
```

### Household Wealth

```{r}
usa_wealth = hh_budget%>%
  filter(Country == 'USA')%>%
  select(Wealth)

usa_wealth%>%
  model(NAIVE(Wealth))%>%
  forecast()%>%
  autoplot(usa_wealth)+
  labs(title = 'Forecast of Wealth in the US')
```

### Australian takeaway food turnover

```{r}
aus_turnover = aus_retail%>%
  filter(Industry == 'Takeaway food services' & State == 'South Australia' )%>%
  select(Turnover)

aus_turnover%>%
  model(SNAIVE(Turnover))%>%
  forecast()%>%
  autoplot(aus_turnover)+
  labs(title = 'South Australian Takeaway Food Turnover')

```

# 5.2

### a)

```{r}
fb = gafa_stock%>%
  filter(Symbol == 'FB')%>%
  update_tsibble(regular=TRUE)%>%
  fill_gaps()%>%
  select(Open)

# plot using stock open price
fb%>%
  ggplot(aes(x = Date, y = Open))+
  geom_line()+
  labs(title = 'Facebook Stock Open Price')
```

### b)

```{r}
fb_forecast = fb%>%
  model(RW(Open~drift()))%>%
  forecast(h=50)
fb_forecast

fb_forecast%>%
  autoplot(fb)+
  labs(title = 'Facebook Stock Price Forecast')
```

### c)

```{r}
end = dim(fb)[1]

# first point 
x_start = fb[1,]$Date
y_start = fb[1,]$Open

# last point
x_end = fb[end,]$Date
y_end = fb[end,]$Open

fb%>%
  model(RW(Open~drift()))%>%
  forecast(h=50)%>%
  autoplot(fb, level=NULL)+
  geom_segment(aes(x = x_start,
                   y = y_start,
                   xend = x_end,
                   yend = y_end))+
  labs(title = 'Facebook Stock Price Forecast')
```

### d)

I think the drift method is the best at forecasting FB stock because it is able to capture the underlying uptrend.

```{r}
fb%>%
  model(NAIVE(Open))%>%
  forecast(h=50)%>%
  autoplot(fb)


fb%>%
  model(SNAIVE(Open~lag()))%>%
  forecast(h=50)%>%
  autoplot(fb)


fb%>%
  model(MEAN(Open))%>%
  forecast(h=50)%>%
  autoplot(fb)
```

# 5.3

From the histogram, we see that the distribution of residuals is almost normal with a mean near 0. This means the SNAIVE model is a good fit as deviations from the model are not correlated

```{r}
# Extract data of interest
recent_production <- aus_production %>%
  filter(year(Quarter) >= 1992)

# Define and estimate a model
fit <- recent_production %>% model(SNAIVE(Beer))
# Look at the residuals
fit %>% gg_tsresiduals()
# Look a some forecasts
fit %>% forecast() %>% autoplot(recent_production)
```

# 5.4

### Australian Exports

As there seems to be no seasonal trends, a NAIVE forecast should work better than SNAIVE for the dataset. Using the NAIVE method, we see that the residuals are normal and not correlated which signal a good fit

```{r}
aus_export = global_economy%>%
  filter(Country == 'Australia')%>%
  select(Exports)

aus_export%>%
  autoplot(Exports)

export_model = aus_export%>%
  model(NAIVE(Exports))
  
export_model%>%
  gg_tsresiduals()

export_model%>%
  forecast()%>%
  autoplot(aus_export)



```

### Brick Production

Looking at the residuals, the NAIVE model seems to be a better fit the australian bricks than the SNAIVE model. This is despite the fact that there seems to be seasonal changes in the original plot

```{r}
aus_brick%>%
  autoplot(Bricks)

brick_model = aus_brick%>%
  model(NAIVE(Bricks))

brick_model%>%
  gg_tsresiduals()

brick_model%>%
  forecast()%>%
  autoplot(aus_brick)
```

# 5.7

### a)

```{r}
set.seed(8009)
myseries <- aus_retail %>%
  filter(`Series ID` == sample(aus_retail$`Series ID`,1))

myseries_train = myseries %>%
  filter(year(Month) < 2011)

head(myseries_train)
```

### b)

Red plot shows less than 2011 filter worked

```{r}
autoplot(myseries, Turnover)+
  autolayer(myseries_train, Turnover, colour = "red")
```

### c)

```{r}
turnover_model = myseries_train %>%
  model(SNAIVE(Turnover))
```

### d)

The residuals are normal with a mean of 0. The variance between residuals seem to have grown larger with time

```{r}
turnover_model%>%
  gg_tsresiduals()
```

### e)

model is able to somewhat accurately capture the future trend

```{r}
turnover_forecast = turnover_model%>%
  forecast(new_data = anti_join(myseries, myseries_train))
turnover_forecast%>%
  autoplot(myseries)
```

### f)

```{r}
turnover_model %>% accuracy()
turnover_forecast %>% accuracy(myseries)
```

### g)

by using less data in our training model, our forecast is less accurate based on all accuracy metrics

```{r}
myseries_train2 = myseries %>%
  filter(year(Month) < 2000)

turnover_forecast2 = myseries_train2%>%
  model(SNAIVE(Turnover))%>%
  forecast(new_data = anti_join(myseries, myseries_train2))

turnover_forecast2%>%
  accuracy(myseries)

turnover_forecast2%>%
  autoplot(myseries)
```

# HW 2 3.3

I attempted to perform box-cox on the dataset from HW 2 3.3 by splitting the time series into two graphs (Increasing and Decreasing)
The results are much better than trying to perform the time series on the entire plot itself

```{r}
get_lambda = function(data, feature){
  data%>%
    features(data[feature],features = guerrero)%>%
    pull(lambda_guerrero)%>%
    return()
}


under_1980 = canadian_gas%>%
  filter_index('1960 Jan' ~ '1980 Jan')

under_1980%>%
  autoplot(box_cox(x = Volume,
                   lambda = get_lambda(under_1980,'Volume')))

over_1980 = canadian_gas%>%
  filter_index('1980 Jan' ~ '2005 Feb')

over_1980%>%
  autoplot(box_cox(x = Volume,
                   lambda = get_lambda(over_1980,'Volume')))

```

