---
title: "data624_hw5"
author: "Richard"
date: "3/13/2022"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
library(fpp3)
library(tidyverse)
```

# 8.1

#### a

```{r}
victorian_pigs = aus_livestock%>%
  filter(State=="Victoria" & Animal=="Pigs")

ETS_pigs = victorian_pigs%>%
  model(ETS(Count~error("A") + trend("A") + season("A")))

ETS_forecast = ETS_pigs%>%
  forecast(h=4)

ETS_forecast%>%
  autoplot(victorian_pigs)+
  geom_line(aes(y= .fitted), col = 'red', data = augment(ETS_pigs))+
  labs(title = 'Number of Pigs Slaughtered in Victoria')+
  guides(colour='none')

```

#### b

intervals are slightly different as the R calculation is based off smoothed values while the manual calculation is on the data before smoothing

```{r}
# mean and variance given
unpack_hilo(hilo(ETS_forecast, 95) , "95%" )%>%
  select(c('95%_lower', '95%_upper'))

# calculated
calc_mean = mean(victorian_pigs$Count)
calc_var = var(victorian_pigs$Count)
c(calc_mean - 1.96*sqrt(calc_var), calc_mean + 1.96*sqrt(calc_var))
```

# 8.5

#### a

US exports have generally gone up over time

```{r}
us_exports = global_economy%>%
  filter(Country == 'United States')%>%
  na.omit()%>%
  select(Exports)

autoplot(us_exports, Exports)+
  labs(title = 'US Exports')
```

#### b

```{r}
ets.ann_export = us_exports%>%
  model(ETS(Exports~error("A") + trend("N") + season("N")))

export_forecast = ets.ann_export%>%
  forecast()

export_forecast%>%
  autoplot(us_exports)+
  geom_line(aes(y= .fitted), col = 'red', 
            data = augment(ets.ann_export))+
  labs(title = 'US Export Forecast')
```

#### c

```{r}
rmse1 = accuracy(ets.ann_export)%>%
  select(RMSE)

rmse1
```

#### d

Since there seems to be an overall trend, ETS(AAN) seems to make sense. However, since there is a short term downtrend at the last point of the dataset, it could be argued that ETS(ANN) is more accurate as it does not include a trend bias
```{r}
ets.aan_export = us_exports%>%
  model(ETS(Exports~error("A") + trend("A") + season("N")))

export_forecast2 = ets.aan_export%>%
  forecast()

export_forecast2%>%
  autoplot(us_exports)+
  geom_line(aes(y= .fitted), col = 'red', 
            data = augment(ets.aan_export))+
  labs(title = 'US Export Forecast')
```

#### e

Based on RMSE, ETS(AAN) performed better than ETS(ANN) on my dataset as the error is slightly lower

```{r}
rmse2 = accuracy(ets.aan_export)%>%
  select(RMSE)

rmse2
```

#### f

Calculated and Given intervals are similar

ETS(ANN)

```{r}
export_mean1 = export_forecast[1,]$.mean

unpack_hilo(hilo(export_forecast, 95) , "95%" )%>%
  select(c('95%_lower', '95%_upper'))

print(c(export_mean1 - 1.96*rmse1$RMSE, 
        export_mean1 + 1.96*rmse1$RMSE))

```

ETS(AAN)

```{r}
export_mean2 = export_forecast2[1,]$.mean

unpack_hilo(hilo(export_forecast2, 95) , "95%" )%>%
  select(c('95%_lower', '95%_upper'))

print(c(export_mean2 - 1.96*rmse2$RMSE, 
        export_mean2 + 1.96*rmse2$RMSE))
```

# 8.6

box-cox ETS is growing exponentially rather than linearly

```{r}
get_lambda = function(data, feature){
  data%>%
    features(data[feature],features = guerrero)%>%
    pull(lambda_guerrero)%>%
    return()
}


chinese_gdp = 
  global_economy%>%
  filter(Country == 'China')%>%
  na.omit()%>%
  select(GDP)

chinese_gdp%>%
  model('AAN' = ETS(GDP~error('A')+ trend('A') + season('N')),
        'AAN (box-cox)' = ETS(box_cox(GDP,get_lambda(chinese_gdp,'GDP'))~error('A')+ trend('A') + season('N')))%>%
  forecast(h=10)%>%
  autoplot(chinese_gdp, level = NULL)+
  labs(title = 'China GDP Forecast')
```

# 8.7

Multiplicative is necessary because there is a trend for the oscillations grow larger with time

The damped model shifts the trend lower, but its hard to tell which is better (multiplicative vs damped) as they are nearly the same


```{r}
aus_gas = aus_production%>%
  select(Gas)

autoplot(aus_gas,Gas)

aus_gas%>%
  model('Holt Multiplication' = ETS(Gas~error('M')+ trend('A') + season('M')),
        'Holt Damped' = ETS(Gas~error('M')+ trend('Ad') + season('M')))%>%
  forecast(h=20)%>%
  autoplot(aus_gas, level=NULL)+
  labs(title = 'Australian gas Forecast')
```



# 8.8

```{r}
set.seed(8009)
myseries <- aus_retail %>%
  filter(`Series ID` == sample(aus_retail$`Series ID`,1))

autoplot(myseries)
```

#### a

Multiplicative method is used because the oscillations grow larger with time (exponential)

#### b

The damped method looks like the forecast is shifted slightly lower. I think the damped version fits this dataset better but, again, it is very similar to the multiplicative method

```{r}
holt_multiply = myseries%>%
  model('Holt Multiplication' = ETS(Turnover~error('M')+ trend('A') + season('M')))
holt_multiply%>%
  forecast(h=20)%>%
  autoplot(myseries, level=NULL)+
  labs(title = 'Holts Multiplicative Forecast')
```
```{r}
holt_damp = myseries%>%
  model('Holt Damped' = ETS(Turnover~error('M')+ trend('Ad') + season('M')))
holt_damp%>%
  forecast(h=20)%>%
  autoplot(myseries, level=NULL)+
  labs(title = 'Holts Damped Forecast')
```

#### c

holts multiplicative method has slightly smaller error than damped
```{r}
accuracy(holt_multiply)%>%
  select(RMSE)
```

```{r}
accuracy(holt_damp)%>%
  select(RMSE)
```

#### e

Based on ACF, not all of the residuals can be considered white noise as some go above the thresholds. However the residuals are normal and centered around 0

```{r}
holt_multiply%>%
  gg_tsresiduals()
```

#### f

Both holts multiplicative and damped models beat the SNAIVE method as they have lower errors

```{r}
snaive = myseries%>%
  model(SNAIVE(Turnover))
snaive%>%
  forecast(h=20)%>%
  autoplot(myseries, level=NULL)+
  labs(title = 'Seasonal Naive Forecast')
```

```{r}
accuracy(snaive)%>%
  select(RMSE)
```

# 8.9 

ETS model on the seasonally adjusted showed the smallest error (RMSE)

```{r}
stl = myseries%>%
  model(STL(box_cox(Turnover,get_lambda(myseries,'Turnover'))))%>%
  components()
stl%>%
  autoplot()
```

```{r}
season_adj = stl%>%
  select(season_adjust)

ets_season = season_adj%>%
  model(ETS(season_adjust~error('A')+ trend('A') + season('N')))

ets_season%>%
  forecast()%>%
  autoplot(season_adj)

```

```{r}
accuracy(ets_season)%>%
  select(RMSE)

```
